---
title: "Developing Data Products Week 4 Project"
author: "Tonya MacDonald"
date: "March 7, 2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This is an RStudio shiny application developed as a part of final project in the Developing Data Products course in Coursera Data Science Specialization track. The application summarizes student data. 

## The Application includes the following:

- Form of Input used: Checkbox
- Reactive output displayed as a result of server logic - different columns will display based on if the checkbox is selected
- The code can be found on github at: <https://github.com/tonyalenore/datasciencecoursera/DevelopingDataProducts>
- The App can be found at: <https://tonyalenore.shinyapps.io/Week4/>

## R Code for the Application

```{r, echo = TRUE, warning=FALSE}
library(shiny)
columns <- c("Gender", "Major", "GPA")
Students <- read.csv("Students.csv")
ui <- fluidPage(
  checkboxGroupInput("features", "Show:",columns),
  tableOutput("data")
)
#titlePanel("Student Information")
server <- function(input, output) {
  output$data <- renderTable({
    Students[, c("Student", input$features), drop = FALSE]
  }, rownames = TRUE)
}
```

## Table of Data 

```{r, echo=FALSE, warning=FALSE}
Students <- read.csv("Students.csv")
Students
```

This is the table of data that was used for this application.
