---
title: "Statistical Inference Course Project"
author: "Tonya MacDonald"
output:
  html_document: default
---

### Purpose

Analyze Tooth Growth Dataset. 

### Libraries

```{r DataLoading}
library("data.table")
library("ggplot2")
#library(dplyr)
#data(ToothGrowth)
```

### Basic summary of the data

```{r}
# show what column names are
names(ToothGrowth)
```

```{r}
# show a bit of the data
head(ToothGrowth)
```

```{r}
# show summary of the data
summary(ToothGrowth)
```

### Compare tooth growth

```{r}
# make dose a factor
ToothGrowth$dose<-as.factor(ToothGrowth$dose)

# show box plot
ggplot(aes(x=dose, y=len), data=ToothGrowth) + 
  geom_boxplot(aes(fill=dose)) + 
  xlab("Dose") + 
  ylab("Length") + 
  facet_grid(~ supp) + 
  ggtitle("Tooth Growth")
```

```{r}
# run t-test
t.test(len~supp,data=ToothGrowth)
```

The p-value is ~ 0.06

---

```{r}
# run t-test using dose amounts 0.5 and 1.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5))
t.test(len~dose,data=ToothGrowth_sub)
```

The p-value is ~ 0, and the confidence interval does not include 0

---

```{r}
# run t-test using dose amounts 1.0 and 2.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,1.0))
t.test(len~dose,data=ToothGrowth_sub)
```

The p-value is ~ 0, and the confidence interval does not include 0

---

```{r}
# run t-test using dose amounts 0.5 and 2.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5,2.0))
t.test(len~dose,data=ToothGrowth_sub)
```

The p-value is ~ 0, and the confidence interval does not include 0

---

### Summary

For each of the two sample t-tests, the p-value is ~ 0, and the confidence interval does not cross over 0.  Also the means increase with dose.

We can conclude that as the dose increases, then the tooth growth increases.  Delivery method does not have an effect on the tooth growth.

